### 2015-02-18 Paper.tex
Add templates

\documentclass[11pt, ngerman, fleqn, DIV=15, headinclude]{scrartcl}

\usepackage[bibatend, color]{header}

\hypersetup{
    pdftitle=
}

%\subject{}
\title{}
%\subtitle{}
\author{
    Martin Ueding \\ \small{\href{mailto:mu@martin-ueding.de}{mu@martin-ueding.de}}
}

\begin{document}

\maketitle

%\newpage
%\tableofcontents
%\newpage

\end{document}

% vim: spell spelllang=de tw=79

### 2015-02-22 Paper.tex
Write more, add data

@@
-\documentclass[11pt, ngerman, fleqn, DIV=15, headinclude]{scrartcl}
+\documentclass[11pt, english, fleqn, DIV=15, headinclude]{scrartcl}
 
 \usepackage[bibatend, color]{header}
+
+\usepackage{pdflscape}
+\usepackage[section]{placeins}
 
@@
 
-%\subject{}
-\title{}
+\subject{physics760 Computational Physics}
+\title{Analysis of $\piup$-$\piup$ lattice QCD scattering data}
 %\subtitle{}
@@
 
-%\newpage
-%\tableofcontents
-%\newpage
+\begin{abstract}
+    % TODO
+\end{abstract}
+
+\tableofcontents
+
+\newpage
+
+\section{Data generation}
+
+\subsection{Metropolis algorithm}
+
+\subsection{Correlation functions}
+
+\section{Analysis methods}
+
+Figure~\ref{fig:analysis-flow} shows the data flow in the analysis. This
+section will go through the whole analysis in the order of the flow chart. The
+methods used will be explained along the way when they are needed.
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=\linewidth]{sketches/Zeichnung.pdf}
+    \caption{%
+        Data flow in the analysis.
+    }
+    \label{fig:analysis-flow}
+\end{figure}
+
+\subsection{Import data}
+
+The data that I was given is organized in different ensembles (like A30.32,
+A100.24). For each ensembles, multiple configurations were simulated. The
+output of each configuration is a two-point correlation function $C_\piup(t)$
+and three four-point correlation functions $C_{\piup\piup}^{(i)}(t)$. Those
+three different contractions are combined into a single
+$\piup\piup$-correlation function:
+\[
+    C_{\piup\piup}(t) = C_{\piup\piup}^{(1)}(t) + C_{\piup\piup}^{(2)}(t)
+    - 2 C_{\piup\piup}^{(3)}(t).
+\]
+
+Some of the ensembles had multiple versions of the data in it. I analyzed all
+of them sorted by file name. In the tables, you will find the ensembles
+multiple times, those are the different versions.
+
+The number of configurations in each ensemble is called $N$.
+
+\subsection{Bootstrap}
+
+All error estimation is done with the bootstrap method. The number of bootstrap
+samples $R$ was set to $R = 3N$.
+
+\subsection{Correlated fit}
+
+The computed masses are shown in table~\ref{tab:masses}.
+
+\subsection{Scattering length}
+
+The computed scattering lengths are shown in table~\ref{tab:masses}.
+
+\section{Results}
+
+\begin{table}
+    \centering
+    \begin{tabular}{lSSS}
+        ensemble & {$m_{\piup}$}  & {$m_{\piup\piup}$} & {$a_0$} \\
+        \midrule
+        A100.24  & 0.22238 +- 0.00023  & 0.45125 +- 0.00052 & -1.346 +- 0.029 \\
+        A100.24  & 0.22233 +- 0.00040  & 0.45083 +- 0.00095 & -1.287 +- 0.132 \\
+        A100.24  & 0.22239 +- 0.00024  & 0.45111 +- 0.00053 & -1.316 +- 0.039 \\
+        A30.32   & 0.12416 +- 0.00055  & 0.25135 +- 0.00144 & -0.904 +- 0.273 \\
+        A40.20   & 0.14779 +- 0.00078  & 0.31429 +- 0.00162 & -1.426 +- 0.052 \\
+        A40.24   & 0.14447 +- 0.00052  & 0.29815 +- 0.00109 & -1.255 +- 0.051 \\
+        A40.24   & 0.14453 +- 0.00031  & 0.29842 +- 0.00071 & -1.272 +- 0.051 \\
+        A40.32   & 0.14126 +- 0.00022  & 0.28626 +- 0.00055 & -1.228 +- 0.092 \\
+        A60.24   & 0.17275 +- 0.00052  & 0.35278 +- 0.00126 & -1.194 +- 0.112 \\
+        A60.24   & 0.17279 +- 0.00048  & 0.35404 +- 0.00099 & -1.361 +- 0.091 \\
+        A80.24   & 0.19930 +- 0.00024  & 0.40511 +- 0.00057 & -1.228 +- 0.043 \\
+        B55.32   & 0.15553 +- 0.00022  & 0.31589 +- 0.00055 & -1.676 +- 0.117 \\
+        D45.32   & 0.12047 +- 0.00046  & 0.25057 +- 0.00137 & -2.416 +- 0.205
+    \end{tabular}
+    \caption{%
+        Computed masses from correlation functions.
+    }
+    \label{tab:masses}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{tabular}{lSSSS}
+        ensemble & {$L$} & {$T$} & {$a_0 m_\piup$} & {$m_\piup/f_\piup$} \\
+        \midrule
+        A100.24 & 24 & 48 & -0.2993 +- 0.0066 & 2.77 \\
+        A100.24 & 24 & 48 & -0.2862 +- 0.0292 & 2.77 \\
+        A100.24 & 24 & 48 & -0.2927 +- 0.0088 & 2.77 \\
+        A30.32  & 32 & 64 & -0.1122 +- 0.0339 & 1.86 \\
+        A40.20  & 20 & 48 & -0.2107 +- 0.0077 & 2.11 \\
+        A40.24  & 24 & 48 & -0.1813 +- 0.0074 & 2.03 \\
+        A40.24  & 24 & 48 & -0.1839 +- 0.0074 & 2.03 \\
+        A40.32  & 32 & 64 & -0.1735 +- 0.0131 & 2.06 \\
+        A60.24  & 24 & 48 & -0.2063 +- 0.0194 & 2.32 \\
+        A60.24  & 24 & 48 & -0.2351 +- 0.0157 & 2.32 \\
+        A80.24  & 24 & 48 & -0.2448 +- 0.0087 & 2.55 \\
+        B55.32  & 32 & 64 & -0.2607 +- 0.0182 & 2.34 \\
+        D45.32  & 32 & 64 & -0.2911 +- 0.0248 & 2.49
+    \end{tabular}
+    \caption{%
+        Lattice size of the ensembles together with computed quantities.
+        These data points are also shown in figure~\ref{fig:result}. The pion
+        decay constants are taken from
+        \parencite[table~1]{Knippschild/Pi_Pi_Scattering}.
+    }
+    \label{tab:computed}
+\end{table}
+
+The results from table~\ref{tab:computed} are shown in figure~\ref{fig:result}.
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=\linewidth]{plots/result.pdf}
+    \caption{%
+        Results. Diamond data points are taken from draft paper.
+    }
+    \label{fig:result}
+\end{figure}
 
@@
 
-% vim: spell spelllang=de tw=79
+% vim: spell spelllang=en tw=79

### 2015-02-22 Paper.tex
Add a caption

@@
     \caption{%
-        Data flow in the analysis.
+        Data flow in the analysis. The first is the import of the data, which
+        is covered in section~\ref{sec:import}. The two and four point
+        functions are combined into pairs for each configuration. The data is
+        then folded in half. The bootstrap step is covered in
+        section~\ref{sec:bootstrap}. Then a correlated fit is performed, see
+        section~\ref{sec:correlated_fit} for details. The second to last step
+        is the application of Lüscher's formula, see
+        section~\ref{sec:scattering_length}. The results are shown in
+        section~\ref{sec:results}.
     }
@@
 \subsection{Import data}
+\label{sec:import}
 
@@
 
+% TODO Folding
+
 \subsection{Bootstrap}
+\label{sec:bootstrap}
 
@@
 \subsection{Correlated fit}
+\label{sec:correlated_fit}
 
@@
 \subsection{Scattering length}
+\label{sec:scattering_length}
 
@@
 \section{Results}
+\label{sec:results}
 

### 2015-02-25 Paper.tex
More sources
Remove spacing, number equation
Write more text in paper

@@
 }
+
+\newcommand\mpi{m_{\piup}}
+\newcommand\mpipi{m_{\piup\piup}}
 
@@
     \caption{%
-        Data flow in the analysis. The first is the import of the data, which
-        is covered in section~\ref{sec:import}. The two and four point
+        Data flow in the analysis. The first step is the import of the data,
+        which is covered in section~\ref{sec:import}. The two and four point
         functions are combined into pairs for each configuration. The data is
@@
 $\piup\piup$-correlation function:
-\[
+\begin{equation}
     C_{\piup\piup}(t) = C_{\piup\piup}^{(1)}(t) + C_{\piup\piup}^{(2)}(t)
     - 2 C_{\piup\piup}^{(3)}(t).
-\]
+\end{equation}
 
@@
 
-% TODO Folding
+In both space and time, the lattice used in the computation has periodic
+boundary conditions. This leads to a symmetry which $t$ and $T-t$ closely
+related. The correlation functions $C(t_1, t_2)$ are symmetric in $t_1$ and
+$t_2$, such that only the data in the interval $[0, T/2]$ carries independent
+information. Before starting any other calculations with the data, it is folded
+in half, like also shown in figure~\ref{fig:folded} later on.
 
@@
 
-The computed masses are shown in table~\ref{tab:masses}.
+The pion masses are contained in the correlation functions in the form of an
+exponential decay constant. Since time is periodic in this simulation, the
+correlation function is also periodic with the temporal lattice extent $T$.
+Therefore, the expectation is not a simple exponential decay but a cosh-like
+function. The four-point functions also have a constant contribution due to the
+finite lattice extent. For both two and four point functions, the functions
+that I fitted to the folded data are:
+\begin{equation}
+    \label{eq:fit2}
+    m_\piup(t, \lambda) = \lambda_1 \sbr{\exp(-\lambda_2 t) + \exp(-\lambda_2
+    [T - t])}
+\end{equation}
+and
+\begin{equation}
+    \label{eq:fit4}
+    m_{\piup\piup}(t, \lambda) = \lambda_1 \sbr{\exp(-\lambda_2 t) + \exp(-\lambda_2
+    [T - t])} + \lambda_3.
+\end{equation}
+
+The folded correlation functions with the fit functions are shown in
+figure~\ref{fig:folded}.
+
+\begin{figure}[htbp]
+    \centering
+    \begin{minipage}[c]{0.49\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{plots/A100_24_L24_T48_beta190_mul0100_musig150_mudel190_kappa1632550__ev120__TB2_SO_LI6_new_c2_folded.pdf}
+    \end{minipage}
+    \begin{minipage}[c]{0.49\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{plots/A100_24_L24_T48_beta190_mul0100_musig150_mudel190_kappa1632550__ev120__TB2_SO_LI6_new_c4_folded.pdf}
+    \end{minipage}
+    \caption{%
+        Folded correlation functions with cosh-like fit function. On the left
+        side, the two point function is shown, the four point function on the
+        right. Only the blue data points were used for fitting. The right
+        ordinate shows the residuals in great magnification (red). One can see
+        that the errors seem way too large. This is a result of the
+        autocorrelation of the data points.
+    }
+    \label{fig:folded}
+\end{figure}
+
+The correlation functions themselves are highly correlated with respect to
+time. Therefore, a regular least squared fit would give a $\chi^2$ that was way
+too low. The $p$-values usually end up around 1, which does not imply a perfect
+fit but that the data fits the model too \emph{well}.
+
+For the correlated fit, a new likelihood function is needed. I chose to keep
+the least squared likelihood function but incorporate the correlation into the
+$\chi^2$. First, a correlation matrix $\tens C$ is needed, which is computed
+from the $R$ bootstrap samples:
+\begin{equation}
+    \label{eq:correlation_matrix}
+    C_{ij} := \frac{1}{R[R-1]} \sum_{k=1}^R
+    [x_{ik} - \bar x_{iR}] [x_{jk} - \bar x_{jR}]
+    \eqnsep
+    \bar x_{iR} := \frac 1R \sum_{r=1}^R x_{ir}.
+\end{equation}
+
+Using this correlation matrix, a new $\chi^2$ can be defined which incorporates
+the inverse correlation matrix:
+\begin{equation}
+    \label{eq:chisq}
+    \chi^2_\text{corr} := \sum_{i, j}^T
+    \left[ \bar x_{iR} - f(t_i, \lambda) \right]
+    C^{-1}_{ij}
+    \left[ \bar x_{jR} - f(t_j, \lambda) \right].
+\end{equation}
+The regular $\chi^2$ has $\tens C^{-1} = \tens 1$ and is just the sum of the
+squared residuals.
+
+In the curve fitting process, a function like
+\texttt{scipy.optimize.curve\_fit} tries to find the parameters $\lambda$ such
+that $\chi^2$ becomes minimal. In my implementation, I used
+\texttt{scipy.optimize.leastsq} which tries to minimize the squared norm of a
+vector valued function. With a Cholesky decomposition, this can be done like
+so: Define the square bracket in eq.~\eqref{eq:chisq} to be the residual vector
+$\vec r$. Then the matrix multiplication can be written as
+\begin{equation}
+    \chi^2 = \vec r^\text T \tens C^{-1} \vec r.
+\end{equation}
+The Cholesky decomposition gives me $\tens C^{-1} = \tens U^\dagger \tens U$
+where $\tens U$ is a upper triangle matrix. Then I can write
+\begin{equation}
+    \chi^2 = [\tens U \vec r]^\dagger [\tens U \vec r]
+    = \| \tens U \vec r \|^2
+\end{equation}
+where $\tens U \vec r$ can be fed into \texttt{scipy.optimize.leastsq}.
+
+As shown in figure~\ref{fig:analysis-flow} the correlation matrix computed from
+the bootstrap samples is also used to fit the original data. Each fit yields a
+value for $\mpi$ and $\mpipi$. The computed masses are shown in
+table~\ref{tab:masses}.
 
@@
 \label{sec:scattering_length}
+
+\parencite{luescher/volume_dependence}:
 
@@
         Lattice size of the ensembles together with computed quantities.
-        These data points are also shown in figure~\ref{fig:result}. The pion
+        These data points are also shown in figure~\ref{fig:result}.
+        The pion
         decay constants are taken from

### 2015-03-13 Paper.tex
Insert hfill
More text
Work on paper and presentation

@@
         Data flow in the analysis. The first step is the import of the data,
-        which is covered in section~\ref{sec:import}. The two and four point
+        which is covered in Section~\ref{sec:import}. The two and four point
         functions are combined into pairs for each configuration. The data is
         then folded in half. The bootstrap step is covered in
-        section~\ref{sec:bootstrap}. Then a correlated fit is performed, see
-        section~\ref{sec:correlated_fit} for details. The second to last step
+        Section~\ref{sec:bootstrap}. Then a correlated fit is performed, see
+        Section~\ref{sec:correlated_fit} for details. The second to last step
         is the application of Lüscher's formula, see
-        section~\ref{sec:scattering_length}. The results are shown in
-        section~\ref{sec:results}.
+        Section~\ref{sec:scattering_length}. The results are shown in
+        Section~\ref{sec:results}.
     }
@@
 The data that I was given is organized in different ensembles (like A30.32,
-A100.24). For each ensembles, multiple configurations were simulated. The
-output of each configuration is a two-point correlation function $C_\piup(t)$
+A100.24). For each ensemble, multiple configurations were simulated. The
+output of each configuration is one two-point correlation function $C_\piup(t)$
 and three four-point correlation functions $C_{\piup\piup}^{(i)}(t)$. Those
@@
 In both space and time, the lattice used in the computation has periodic
-boundary conditions. This leads to a symmetry which $t$ and $T-t$ closely
-related. The correlation functions $C(t_1, t_2)$ are symmetric in $t_1$ and
-$t_2$, such that only the data in the interval $[0, T/2]$ carries independent
-information. Before starting any other calculations with the data, it is folded
-in half, like also shown in figure~\ref{fig:folded} later on.
+boundary conditions. The spatial lattice extent is called $L$, the temporal one
+$T$. This leads to a symmetry where the time slices $t$ and $T-t$ closely
+related. The correlation functions $C(t)$ with $t = |t_1 - t_2|$ are symmetric
+in $t_1$ and $t_2$, such that only the data in the interval $[0, T/2]$ carries
+independent information. Before starting any other calculations with the data,
+it is folded in half and averaged. The data then looks like in
+Figure~\ref{fig:folded} which is shown later on.
 
@@
 
-All error estimation is done with the bootstrap method. The number of bootstrap
-samples $R$ was set to $R = 3N$.
+All error estimation is done with the bootstrap method. One other known method
+to compute errors is the Gaussian error propagation. This method involves
+computing derivatives of the functions applied to the data, which can be either
+numerically unstable or just not really applicable: How does the variance of
+the mean change when you change the input data?
+
+In order to get rid of all those problems at the same time one does a meta
+analysis of the data itself. To get back to the example of the mean, let $X$ be
+the set of $N$ data points ($x_i \in \R$). Then the estimator of the mean
+$\bar\mu$ can be computed directly from $X$. Now $R$ bootstrap samples $X^r$
+are computed from the set by randomly selecting $N$ elements from $X$ and
+putting them into $X^r$ (upper index, no power). Since $X$ is sampled from the
+original distribution, it resembles the original distribution as good as it can
+with only $N$ elements. $X^r$ therefore is sampled from an approximation of the
+original distribution. Then the estimators for the means $\bar\mu^r$ are
+computed from the bootstrapped sets. The mean of all the $\bar\mu^r$ should
+give $\bar\mu$, which was computed from the original data. The error estimate
+is then the standard deviation of all the $\bar\mu^r$.
+
+The number of bootstrap samples $R$ was set to $R = 3N$ for each ensemble.
 
@@
 correlation function is also periodic with the temporal lattice extent $T$.
-Therefore, the expectation is not a simple exponential decay but a cosh-like
-function. The four-point functions also have a constant contribution due to the
-finite lattice extent. For both two and four point functions, the functions
-that I fitted to the folded data are:
+Therefore, the expectation is not a simple exponential decay—$\exp(-\lambda
+t)$—but a cosh-like function. The four-point functions also have a constant
+contribution due to the finite lattice extent. For both two and four point
+functions, the functions that I fitted to the folded data with parameters $\vec
+\lambda$ are:
 \begin{equation}
@@
 The folded correlation functions with the fit functions are shown in
-figure~\ref{fig:folded}.
+Figure~\ref{fig:folded}.
 
@@
     \end{minipage}
+    \hfill
     \begin{minipage}[c]{0.49\linewidth}
@@
     \caption{%
-        Folded correlation functions with cosh-like fit function. On the left
-        side, the two point function is shown, the four point function on the
-        right. Only the blue data points were used for fitting. The right
-        ordinate shows the residuals in great magnification (red). One can see
-        that the errors seem way too large. This is a result of the
-        autocorrelation of the data points.
+        Folded correlation functions with cosh-like fit function using a
+        logarithmic ordinate scale. On the left side, the two point function is
+        shown, the four point function on the right. Only the blue data points
+        were used for fitting. The right ordinate shows the residuals in great
+        magnification (red). One can see that the errors seem way too large.
+        This is a result of the autocorrelation of the data points.
     }
@@
 The correlation functions themselves are highly correlated with respect to
-time. Therefore, a regular least squared fit would give a $\chi^2$ that was way
-too low. The $p$-values usually end up around 1, which does not imply a perfect
-fit but that the data fits the model too \emph{well}.
+time. Therefore, a regular least squared fit would give a $\chi^2$ that would
+be way too low. The $p$-values usually end up around 1, which does not imply a
+perfect fit but rather that the data fits the model too \emph{well}. This, in
+turn, means that errors and residuals are over-estimated with the given model.
+A different model is called for: the correlated fit.
 
@@
 vector valued function. With a Cholesky decomposition, this can be done like
-so: Define the square bracket in eq.~\eqref{eq:chisq} to be the residual vector
+so: Define the square bracket in Eq.~\eqref{eq:chisq} to be the residual vector
 $\vec r$. Then the matrix multiplication can be written as
@@
 \end{equation}
-The Cholesky decomposition gives me $\tens C^{-1} = \tens U^\dagger \tens U$
+The Cholesky decomposition gives me $\tens C^{-1} = \tens U^\dagger \tens U$,
 where $\tens U$ is a upper triangle matrix. Then I can write
@@
 \end{equation}
-where $\tens U \vec r$ can be fed into \texttt{scipy.optimize.leastsq}.
+where $\tens U \vec r$ is a vector that can be fed into
+\texttt{scipy.optimize.leastsq}.
 
-As shown in figure~\ref{fig:analysis-flow} the correlation matrix computed from
+As shown in Figure~\ref{fig:analysis-flow}, the correlation matrix computed from
 the bootstrap samples is also used to fit the original data. Each fit yields a
 value for $\mpi$ and $\mpipi$. The computed masses are shown in
-table~\ref{tab:masses}.
+Table~\ref{tab:masses}.
 
@@
 
-\parencite{luescher/volume_dependence}:
+There is a relation between mass difference to scattering length
+\parencite[(1.3)]{luescher/volume_dependence}:
+\begin{equation}
+    \label{eq:luescher}
+    \mpipi = 2\mpi - \frac{4 \piup a_0}{\mpi L^3} \sbr{1 + c_1 \frac{a_0}L + c_2 \frac{a_0^2}{L^2}}
+    \eqnsep
+    c_1 = \num{-2.837297}
+    \eqnsep
+    c_2 = \num{6.375183}
+\end{equation}
 
-The computed scattering lengths are shown in table~\ref{tab:masses}.
+Using the computed $\mpi$ and $\mpipi$ with Lüscher's formula, I can
+numerically solve for the scattering length $a_0$. In my program, I use
+\texttt{scipy.optimize.brentq} that is based on the algorithm by Brent (1973).
+
+The computed scattering lengths are shown in Table~\ref{tab:masses}.
 
@@
     \caption{%
-        Computed masses from correlation functions.
+        Computed masses from correlation functions. The last column shows the
+        scattering length $a_0$ which is computed using Lüscher's formula,
+        Eq.~\eqref{eq:luescher}.
     }
@@
         Lattice size of the ensembles together with computed quantities.
-        These data points are also shown in figure~\ref{fig:result}.
+        These data points are also shown in Figure~\ref{fig:result}.
         The pion
@@
 
-The results from table~\ref{tab:computed} are shown in figure~\ref{fig:result}.
+The results from Table~\ref{tab:computed} are shown in Figure~\ref{fig:result}.
 
@@
     \caption{%
-        Results. Diamond data points are taken from draft paper.
+        $\mpi a_0$ as a function of $\mpi/f_\piup$ as shown in
+        Table~\ref{tab:computed}. Diamond data points are taken from draft
+        paper from Carsten Urbach's workgroup. Some ensembles had multiple
+        versions of the correlation functions in them, this plot shows all of
+        them in lexical order of the pathname.
     }

### 2015-03-16 Paper.tex
Write a bit more

@@
-\documentclass[11pt, english, fleqn, DIV=15, headinclude]{scrartcl}
+\documentclass[11pt, english, fleqn, DIV=10, headinclude]{scrartcl}
 
@@
 \begin{abstract}
-    % TODO
+    Correlation functions of pions that were computed with lattice QCD methods
+    are analyzed. The effective masses of single pions and two interacting
+    pions are extracted. Using a finite size formula by Lüscher, the S-wave
+    scattering length is computed.
 \end{abstract}
@@
 
-\subsection{Metropolis algorithm}
+% Weighted generation with HMC
 
-\subsection{Correlation functions}
+% Downloadable ensembles
+
+% Operators on data
+
+% Contractions
 
@@
 $\bar\mu$ can be computed directly from $X$. Now $R$ bootstrap samples $X^r$
-are computed from the set by randomly selecting $N$ elements from $X$ and
-putting them into $X^r$ (upper index, no power). Since $X$ is sampled from the
+(upper index, no power) are computed from the set by randomly selecting $N$
+elements from $X$ and putting them into $X^r$. Since $X$ is sampled from the
 original distribution, it resembles the original distribution as good as it can
@@
 correlation function is also periodic with the temporal lattice extent $T$.
-Therefore, the expectation is not a simple exponential decay—$\exp(-\lambda
-t)$—but a cosh-like function. The four-point functions also have a constant
+Therefore, the expectation is not a simple exponential decay, $\exp(-\lambda
+t)$, but a cosh-like function. The four-point functions also have a constant
 contribution due to the finite lattice extent. For both two and four point
@@
     \caption{%
-        Folded correlation functions with cosh-like fit function using a
-        logarithmic ordinate scale. On the left side, the two point function is
-        shown, the four point function on the right. Only the blue data points
-        were used for fitting. The right ordinate shows the residuals in great
-        magnification (red). One can see that the errors seem way too large.
-        This is a result of the autocorrelation of the data points.
+        Folded correlation functions with cosh-like fit function. On the left
+        side, the two point function is shown using a logarithmic ordinate
+        scale. The four point function is shown on the right. Only the blue
+        data points were used for fitting. The right ordinate shows the
+        residuals in great magnification (red). One can see that the errors
+        seem way too large compared to the residuals. This is a result of the
+        autocorrelation of the data points.
     }
@@
 time. Therefore, a regular least squared fit would give a $\chi^2$ that would
-be way too low. The $p$-values usually end up around 1, which does not imply a
-perfect fit but rather that the data fits the model too \emph{well}. This, in
-turn, means that errors and residuals are over-estimated with the given model.
-A different model is called for: the correlated fit.
+be way too low for the assumed degrees of freedom. The $p$-values usually end
+up around 1, which does not imply a perfect fit but rather that the data fits
+the model \emph{too well}. This, in turn, means that errors and residuals are
+over-estimated with the given model. A different model is called for: the
+correlated fit.
 
@@
     \label{eq:correlation_matrix}
-    C_{ij} := \frac{1}{R[R-1]} \sum_{k=1}^R
-    [x_{ik} - \bar x_{iR}] [x_{jk} - \bar x_{jR}]
+    C_{ij} := \frac{1}{R[R-1]} \sum_{r=1}^R
+    [x_{ir} - \bar x_{iR}] [x_{jr} - \bar x_{jR}]
     \eqnsep
@@
 vector valued function. With a Cholesky decomposition, this can be done like
-so: Define the square bracket in Eq.~\eqref{eq:chisq} to be the residual vector
-$\vec r$. Then the matrix multiplication can be written as
+so: Define the square bracket in Equation~\eqref{eq:chisq} to be the residual
+vector $\vec r$. Then the matrix multiplication can be written as
 \begin{equation}
@@
 \texttt{scipy.optimize.brentq} that is based on the algorithm by Brent (1973).
+The computed scattering lengths are shown in Table~\ref{tab:masses}.
 
-The computed scattering lengths are shown in Table~\ref{tab:masses}.
+Equation~\eqref{eq:luescher} can be motivated like so: Let $\hat H = \hat H_0 +
+\hat V$ be the Hamiltonian for two identical particles where $\hat H_0$ is the
+free Hamiltonian and $\hat V$ a spherically symmetric potential. The
+eigenfunctions of $\hat H_0$ are plane waves which are denoted $\ket{\vec p,
+\vec q}$. In continous space, the normalization is
+\parencite[(2.10)]{luescher/volume_dependence}
+\[
+    \braket{\vec p', \vec q'|\vec p, \vec q}
+    = [2\piup]^6 \sbr{
+        \delta(\vec p' - \vec p)
+        \delta(\vec q' - \vec q)
+        +
+        \delta(\vec p' - \vec q)
+        \delta(\vec q' - \vec p)
+    }.
+\]
+On the discrete lattice, the possible momentum values are restricted to lattice
+values and the potential will be periodic due to periodic boundary conditions.
+This changes the normalization to
+\parencite[(2.21)]{luescher/volume_dependence}
+\[
+    \braket{\vec p', \vec q'|\vec p, \vec q}
+    = L^6 \sbr{
+        \delta_{\vec p', \vec p}
+        \delta_{\vec q', \vec q}
+        +
+        \delta_{\vec p', \vec q}
+        \delta_{\vec q', \vec p}
+    }.
+\]
+
+The matrix element of the potential $\vec V$ are given by
+\parencite[(2.21)]{luescher/volume_dependence}
+\[
+    \braket{\vec p', \vec q'| \hat V | \vec p, \vec q}
+    = L^3 \delta_{\vec P', \vec P} \hat V\del{\frac12 [\vec p' - \vec q'],
+    \frac12 [\vec p - \vec q]},
+\]
+where $\vec P$ is the total momentum of the two particles. The energy shift in
+first order of $V$ for the ground state (all momenta are zero) then is the
+quotient \parencite[(2.24)]{luescher/volume_dependence}:
+\[
+    \Deltaup E = \frac{1}{2L^3} \hat V(\vec 0, \vec 0) + \mathcal O(V^2).
+\]
+The scattering amplitude as well as the S-wave scattering length $a_0$ can be
+expanded in powers of $\hat V$ (Born series). Using first order term from 
+\parencite[(2.18)]{luescher/volume_dependence}, above equation can be rewritten
+with the scattering length
+\parencite[(2.25)]{luescher/volume_dependence}:
+\[
+    \Deltaup E = - \frac{4\piup a_0}{m L^3} + \mathcal O(V^2).
+\]
+
+To first order on $V$, the leading term scales with $L^{-3}$, like in
+Equation~\eqref{eq:luescher}. To obtain more terms, a complete perturbation
+expansion needs to be written out. That expansion will be both in terms of
+$L^{-1}$ \emph{and} $V$. \citeauthor{luescher/volume_dependence} then was able
+to group those terms in powers of $L^{-1}$ and obtain
+Equation~\eqref{eq:luescher}.
+
+% FIXME Maybe write a bit more on the methods?
 
@@
 \label{sec:results}
+
+My analysis of the correlation functions gives me values for the masses of a
+single pion ($\mpi$) and a pair of them ($\mpipi$). Using
+Equation~\eqref{eq:luescher}, I was able to compute the S-wave scattering
+length $a_0$ from the energy difference. Those values are listed in
+Table~\ref{tab:masses}.
 
@@
         scattering length $a_0$ which is computed using Lüscher's formula,
-        Eq.~\eqref{eq:luescher}.
+        Equation~\eqref{eq:luescher}.
     }
@@
 \end{table}
+
+The different ensembles assume different masses for the quarks. The first
+number behind the letter gives the assumed quark mass. In most cases, a lower
+quark mass also yields a lower pion mass $\mpi$. Those pion masses are not
+directly the physical ones, one needs to extrapolate to the physical point. The
+masses and scattering lengths are given in lattice units and depends on the
+lattice spacing. The product of mass and scattering length, $a_0 \mpi$, is
+independent of the lattice spacing and a good measure for extrapolation.
+Table~\ref{tab:computed} shows this product, as well as the decay constant.
 
@@
         Lattice size of the ensembles together with computed quantities.
-        These data points are also shown in Figure~\ref{fig:result}.
-        The pion
+        These data points are also shown in Figure~\ref{fig:result}. The pion
         decay constants are taken from

### 2015-03-17 Paper.tex
Write something about the data generation

@@
 
-% Weighted generation with HMC
+Monte Carlo methods can be used to simulate quantum systems. Key pieces are
+Feynman's path integral formalism and Euclidean spacetime where the metric
+tensor $\tens\eta$ has signature $\pm 4$ and the weight factor in the path
+integral is $\exp(-S_\text{cl})$ instead of $\exp(\iup S_\text{cl})$ where
+$S_\text{cl}$ is the classical action
+\parencite[Section~2]{Creutz/Statistical_Approach_QM}. A Monte Carlo algorithm
+like the Metropolis algorithm can then be used to generate configurations which
+are already distributed by the weight of the classical action. Therefore,
+expectational values of observables are simply calculated and averaged over all
+available configurations \parencite[(3.7)]{Creutz/Statistical_Approach_QM}.
 
-% Downloadable ensembles
+The group of Andreas Kell, Martin Efferz and Simon Blanke have used this for
+the harmonic oscillator in their Computational Physics project this year. I
+also did this in my bachelor's thesis: \parencite{Ueding/Bachelorarbeit}.
 
-% Operators on data
+For quantum chromodynamics (QCD), the theory of the strong force, this
+methodology can be used as well. The action of QCD is used with a hybrid Monte
+Carlo algorithm to generate configurations of gauge fields on a four
+dimensional lattice. This process is done for several ensembles which differ in
+parameters like the quark mass. Ensemble generation takes a lot of computing
+time, so the results are made available to download.
+% TODO Would be nice to have a location where this could be downloaded.
 
-% Contractions
+As with the harmonic oscillator on the lattice, expectation values of
+observables can be computed for each configuration of the ensembles and then
+averaged (without weight) over all ensembles yielding an estimate for the
+observable. Since this uses a field theory, particles of interested can be
+created at the beginning of time and annihilated at the end of the temporal
+lattice extent. This way, specific particle configurations (like
+$\piup$–$\piup$ scattering) can be examined without needing to generate new
+configurations.
+
+Using the correct operators, it is possible to construct correlation functions
+which contain information about the energies/masses of the involved particles,
+analogously to \parencite[(4.14)]{Creutz/Statistical_Approach_QM} for the
+harmonic oscillator. I was given those correlation functions to analyze.
 

### 2015-03-17 Paper.tex
Restructure Lüscher formula section

@@
 \subject{physics760 Computational Physics}
-\title{Analysis of $\piup$-$\piup$ lattice QCD scattering data}
+\title{Analysis of $\piup$-$\piup$ lattice QCD \\ scattering data}
 %\subtitle{}
@@
 is then the standard deviation of all the $\bar\mu^r$.
+\parencite[Slide~5]{Oser/Bootstrap}
 
@@
 $\chi^2$. First, a correlation matrix $\tens C$ is needed, which is computed
-from the $R$ bootstrap samples:
+from the $R$ bootstrap samples: \parencite[Section~2]{Michael:1994sz}
 \begin{equation}
@@
 \hat V$ be the Hamiltonian for two identical particles where $\hat H_0$ is the
-free Hamiltonian and $\hat V$ a spherically symmetric potential. The
-eigenfunctions of $\hat H_0$ are plane waves which are denoted $\ket{\vec p,
-\vec q}$. In continous space, the normalization is
-\parencite[(2.10)]{luescher/volume_dependence}
-\[
-    \braket{\vec p', \vec q'|\vec p, \vec q}
-    = [2\piup]^6 \sbr{
-        \delta(\vec p' - \vec p)
-        \delta(\vec q' - \vec q)
-        +
-        \delta(\vec p' - \vec q)
-        \delta(\vec q' - \vec p)
-    }.
-\]
-On the discrete lattice, the possible momentum values are restricted to lattice
-values and the potential will be periodic due to periodic boundary conditions.
-This changes the normalization to
-\parencite[(2.21)]{luescher/volume_dependence}
-\[
-    \braket{\vec p', \vec q'|\vec p, \vec q}
-    = L^6 \sbr{
-        \delta_{\vec p', \vec p}
-        \delta_{\vec q', \vec q}
-        +
-        \delta_{\vec p', \vec q}
-        \delta_{\vec q', \vec p}
-    }.
-\]
+free Hamiltonian and $\hat V$ a spherically symmetric potential with a limited
+range (or faster than $1/\sqrt r$ decay). Then the probability that the two
+particles are in interaction range scales like $L^{-3}$ as this is the inverse
+volume which the two particles occupy. Then the most prominent finite size
+effect scales to the third power, all other effects can only have higher
+powers.
 
-The matrix element of the potential $\vec V$ are given by
-\parencite[(2.21)]{luescher/volume_dependence}
-\[
-    \braket{\vec p', \vec q'| \hat V | \vec p, \vec q}
-    = L^3 \delta_{\vec P', \vec P} \hat V\del{\frac12 [\vec p' - \vec q'],
-    \frac12 [\vec p - \vec q]},
-\]
-where $\vec P$ is the total momentum of the two particles. The energy shift in
-first order of $V$ for the ground state (all momenta are zero) then is the
-quotient \parencite[(2.24)]{luescher/volume_dependence}:
+In the first order of the potential $V$ the ground state (all momenta are zero)
+will have an energy shift proportional to $L^{-3}$
+\parencite[(2.24)]{luescher/volume_dependence}:
 \[
@@
 \]
-The scattering amplitude as well as the S-wave scattering length $a_0$ can be
-expanded in powers of $\hat V$ (Born series). Using first order term from 
+The scattering amplitude $T$ as well as the S-wave scattering length $a_0$ can
+be expanded in powers of $\hat V$ (Born series). Using first order term from
 \parencite[(2.18)]{luescher/volume_dependence}, above equation can be rewritten
@@
 expansion needs to be written out. That expansion will be both in terms of
-$L^{-1}$ \emph{and} $V$. \citeauthor{luescher/volume_dependence} then was able
-to group those terms in powers of $L^{-1}$ and obtain
+$L^{-1}$ and $V$. \citeauthor{luescher/volume_dependence} then was able to
+group those terms in powers of $L^{-1}$ and obtain
 Equation~\eqref{eq:luescher}.
-
-% FIXME Maybe write a bit more on the methods?
 
@@
 The results from Table~\ref{tab:computed} are shown in Figure~\ref{fig:result}.
+My data points have no marker, the reference values have a diamond marker. The
+solid black line is the expectation, not a fit.
+
+As can be seen with the comparison to the reference data, that my values are a
+lot less accurate than the reference ones. For most points, my errors are a lot
+larger. The statistical uncertainty on my values is high enough for them to be
+compatible with the reference and expectation.
 

### 2015-03-18 Paper.tex
Improve small things
Write a tad more

@@
 Monte Carlo methods can be used to simulate quantum systems. Key pieces are
-Feynman's path integral formalism and Euclidean spacetime where the metric
+Feynman's path integral formalism and Euclidean spacetime. In the latter, the metric
 tensor $\tens\eta$ has signature $\pm 4$ and the weight factor in the path
-integral is $\exp(-S_\text{cl})$ instead of $\exp(\iup S_\text{cl})$ where
+integral is $\exp(-S_\text{cl})$ instead of $\exp(\iup S_\text{cl})$; 
 $S_\text{cl}$ is the classical action
@@
 the harmonic oscillator in their Computational Physics project this year. I
-also did this in my bachelor's thesis: \parencite{Ueding/Bachelorarbeit}.
+also did this in my bachelor's thesis last year \parencite{Ueding/Bachelorarbeit}.
 
@@
 averaged (without weight) over all ensembles yielding an estimate for the
-observable. Since this uses a field theory, particles of interested can be
-created at the beginning of time and annihilated at the end of the temporal
-lattice extent. This way, specific particle configurations (like
-$\piup$–$\piup$ scattering) can be examined without needing to generate new
-configurations.
+observable. Since this uses a field theory, particles of interest can be
+created at the beginning of time and annihilated laster on. This way, specific
+particle configurations (like $\piup$–$\piup$ scattering) can be examined
+without needing to generate new configurations.
 
@@
 The data that I was given is organized in different ensembles (like A30.32,
-A100.24). For each ensemble, multiple configurations were simulated. The
-output of each configuration is one two-point correlation function $C_\piup(t)$
-and three four-point correlation functions $C_{\piup\piup}^{(i)}(t)$. Those
-three different contractions are combined into a single
-$\piup\piup$-correlation function:
+A100.24) as mentioned in the introduction. For each ensemble, multiple
+configurations were simulated. The output of each configuration is one
+two-point correlation function $C_\piup(t)$ and three four-point correlation
+functions $C_{\piup\piup}^{(i)}(t)$. Those three different contractions are
+combined into a single $\piup\piup$-correlation function:
 \begin{equation}
@@
 it is folded in half and averaged. The data then looks like in
-Figure~\ref{fig:folded} which is shown later on.
+Figure~\ref{fig:folded} which appears laster on.
 
@@
 \end{equation}
+The parameter $\lambda_3$ is a lattice artifact that will vanish for $L \to
+\infty$.
 
@@
 For the correlated fit, a new likelihood function is needed. I chose to keep
-the least squared likelihood function but incorporate the correlation into the
-$\chi^2$. First, a correlation matrix $\tens C$ is needed, which is computed
-from the $R$ bootstrap samples: \parencite[Section~2]{Michael:1994sz}
+the falling exponential likelihood function but incorporate the correlation
+into the $\chi^2$. First, a correlation matrix $\tens C$ is needed, which is
+computed from the $R$ bootstrap samples: \parencite[Section~2]{Michael:1994sz}
 \begin{equation}
@@
 In the curve fitting process, a function like
-\texttt{scipy.optimize.curve\_fit} tries to find the parameters $\lambda$ such
-that $\chi^2$ becomes minimal. In my implementation, I used
+\texttt{scipy.optimize.curve\_fit} tries to find the parameters $\vec\lambda$
+such that $\chi^2$ becomes minimal. In my implementation, I used the function
 \texttt{scipy.optimize.leastsq} which tries to minimize the squared norm of a
@@
 so: Define the square bracket in Equation~\eqref{eq:chisq} to be the residual
-vector $\vec r$. Then the matrix multiplication can be written as
+vector $\vec r(\vec \lambda)$. Then the matrix multiplication can be written as
 \begin{equation}
@@
 \texttt{scipy.optimize.leastsq}.
+\nocite{SciPy}
 
@@
 particles are in interaction range scales like $L^{-3}$ as this is the inverse
-volume which the two particles occupy. Then the most prominent finite size
-effect scales to the third power, all other effects can only have higher
-powers.
+volume which the two particles occupy. The most prominent finite size effect
+scales to the third power, all other effects can only have higher powers of
+$L^{-1}$.
 
-In the first order of the potential $V$ the ground state (all momenta are zero)
+In the first order of the potential $V$, the ground state (all momenta are zero)
 will have an energy shift proportional to $L^{-3}$
@@
 
-To first order on $V$, the leading term scales with $L^{-3}$, like in
-Equation~\eqref{eq:luescher}. To obtain more terms, a complete perturbation
-expansion needs to be written out. That expansion will be both in terms of
-$L^{-1}$ and $V$. \citeauthor{luescher/volume_dependence} then was able to
-group those terms in powers of $L^{-1}$ and obtain
+To first order on the potential $V$, the leading term scales with $L^{-3}$,
+like in Equation~\eqref{eq:luescher}. To obtain more terms, a complete
+perturbation expansion needs to be written out. That expansion will be both in
+terms of $L^{-1}$ and $V$. \citeauthor{luescher/volume_dependence} then was
+able to group those terms in powers of $L^{-1}$ and obtain
 Equation~\eqref{eq:luescher}.
@@
 
+The range of fitting was set manually by looking at the plateau of the
+effective masses. This introduces some statistical errors which cannot be
+quantified with the current machinery. A way to get beyond this limitation is
+to run the whole analysis with all possible fit ranges. The different result
+are then weighted by the $p$-values of the respective fit.
+
+\clearpage
+
 \end{document}

### 2015-03-18 Paper.tex
Add effective masses plot

@@
 
-The folded correlation functions with the fit functions are shown in
-Figure~\ref{fig:folded}.
+In order to determine the fit range, I have used the effective mass as shown in
+Figure~\ref{fig:m_eff}. The plateau starts at $t/a = 13$, so those are the
+points I have used. The folded correlation functions with the fit functions are
+shown in Figure~\ref{fig:folded}.
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=\linewidth]{plots/A100_24_L24_T48_beta190_mul0100_musig150_mudel190_kappa1632550__ev120__TB2_SO_LI6_new_c2_m_eff.pdf}
+    \caption{%
+        Effective mass of A100.24 ensemble.
+    }
+    \label{fig:m_eff}
+\end{figure}
 
@@
         seem way too large compared to the residuals. This is a result of the
-        autocorrelation of the data points.
+        autocorrelation of the data points. Ensemble is A100.24.
     }
@@
 The range of fitting was set manually by looking at the plateau of the
-effective masses. This introduces some statistical errors which cannot be
-quantified with the current machinery. A way to get beyond this limitation is
-to run the whole analysis with all possible fit ranges. The different result
-are then weighted by the $p$-values of the respective fit.
+effective masses (see Figure~\ref{fig:m_eff}). This introduces some statistical
+errors which cannot be quantified with the current machinery. A way to get
+beyond this limitation is to run the whole analysis with all possible fit
+ranges. The different result are then weighted by the $p$-values of the
+respective fit.
 
